---
title: "lab-07-simpsons.Rmd"
author: "Asrar Almutairi "
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham
```
    
Your answer: 
the data is observational as description states that is based on age, and mortality which are all observational events and not produced via experiment.

2.  
```{r observations}
nrow(Whickham)
```   

Your answer; 
There are 1,314 obs. As we Know every row is an obs

3.  

```{r number_variables}
names(Whickham)
```

Your answer:  
There are 3 variables, "outcome", "smoker", "age".

```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)
unique(Whickham$age)
```

Your answer: 
Using the "unique()" function on the 3 variables we could see that "outcome" only takes Alive or Dead vslue, which makes it categorical non-ordinal. "smoker" only takes Yes or No, which also makes it categorical non-ordinal. Age is numerical continous data.

One of the best ways to visualise categorical data is through the use of bar charts.
```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +
  geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +
  geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age)) +
  geom_boxplot()
```


4.
Smoking is a bad habit if it continues. It can lead to death and increase the proportion.

```{r plot_smokeVSincome_1}
ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")
```

Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)
```

6.

```{r age_cat}

```

7. 

```{r plot_smokeVSincome_3}

```

Knit, commit, and push to github.