---
title: "lab-07-simpsons.Rmd"
author: "Asrar Almutairi "
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham
```
    
Your answer: 
the data is observational as description states that is based on age, and mortality which are all observational events and not produced via experiment.

2.  
```{r observations}
nrow(Whickham)
```   

Your answer; 
There are 1,314 obs. As we Know every row is an obs

3.  

```{r number_variables}
names(Whickham)
```

Your answer:  
There are 3 variables, "outcome", "smoker", "age".

```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)
unique(Whickham$age)
```

Your answer: 
Using the "unique()" function on the 3 variables we could see that "outcome" only takes Alive or Dead vslue, which makes it categorical non-ordinal. "smoker" only takes Yes or No, which also makes it categorical non-ordinal. Age is numerical continous data.

One of the best ways to visualise categorical data is through the use of bar charts.
```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +
  geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +
  geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age)) +
  geom_boxplot()
```


4.
Smoking is a bad habit if it continues. It can lead to death and increase the proportion.

```{r plot_smokeVSincome_1}
ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")
```

Knit, commit, and push to github.

5. 
smoker no  (732) :   31,4 (dead) >> (68,6) alive
smoker yes (582) :   23,8 (dead) >> (76,2) alive

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)
502+230
230/732
```

6.

```{r age_cat}
Whickham <- Whickham%>% mutate (age_cat = case_when (age <= 44 ~ "18-44", age > 44. & age <= 64 ~ "45-64", age > 64 ~ "65+"))
```

7. 
what changes > the category of the age it's appear to us and we see the most of dead people not smoker in age (65+).. but in age (45-64)and (18-44)the most dead people are smoker that is relationship between the smoking and helth not clearly but can say that your helth will be change to worst if you be smoker.

```{r plot_smokeVSincome_3}
ggplot(data=Whickham, aes(x=outcome, y=smoker,color=smoker, fill=age_cat)) + geom_bar(stat="identity") +facet_wrap(~age_cat)
```

Knit, commit, and push to github.